---
title: "Moore-Data-Analysis"
format: html
editor: visual
---

Setting up workspace

```{r}
#Clearing R's environment
rm (list = ls())
```

```{r}
#Activating necessary packages
library(tidyverse)
library(ggfortify)
library(here)
```

```{r}
#Reading in data
eastwood <- read.csv(here("Data", "Eastwood23_superb_climate_data_revised.csv"))
```

#Question 1: Is there a relationship between rainfall amounts and telomere length? I am looking for a relationship between rainfall and telomere length, which are both ratio data. Since there is only one predictor variable, I will use a linear regression.

Predicting analysis

```{r}
#Making scatterplot as initial depiction of data. Making predictions off of this scatterplot can allow me to determine if there are any mistakes when I do my analysis. The summary statistics in the analysis should be similar to the ones estimated from the scatterplot.
ggplot(eastwood, aes(Rainfall.nestling, rTL)) + geom_point()

#The telomere lengths range between 0.25-1.20, and the rainfall amounts range between 0-95. The slope should be about (1.2-.25)/(95-0) = 0.01. The y-intercept appears to be around 0.25. Because of this low slope, I do not expect there to be a significant relationship between telomere length and rainfall amounts. Also, there does not appear to be a relationship based just upon a visual evaluation of the graph.
```

Making linear regression

```{r}
model_rainfall <- lm(rTL ~ Rainfall.nestling, data = eastwood)
```

Checking assumption

```{r}
#The residuals vs fitted plot should have no humps or valleys to assume that a linear model is appropriate for this data. In the Q-Q plot, the data should be linear to assume that the data is normally distributed. In the scale-location plot, there should be no pattern of points, so the line should be straight, to assume that there is constant variance. The residuals vs leverage plot detects points that have too much influence on the data set, such as outliers. 
autoplot(model_rainfall, smooth.color = NA)

#The Q-Q plot looks good, but the residuals vs fitted, scale-location, and residuals vs leverage plots are not as linear as they should be. There are 4 points that might be skewing this data, but the reason that I did not remove them before was because these outliers did not show up on a boxplot in my data exploration stage (seperate file). I will continue with my analysis as if the data fits the model assumptions, but later, should I remove these?
```

Running ANOVA

```{r}
#ANOVA test
anova(model_rainfall)
#The F-statistic is 0.125 and the p-value is 0.7236, which does not indicate a significant relationship between rainfall amount and telomere length. The degrees of freedom + residuals + 1 should add up to the number of rows, which is the case because 1 + 341 + 1 = 343. 

#Viewing results in a summary table
summary(model_rainfall)
#The adjusted R-squared is -0.002564. This value indicates that the variation in telomere length is not highly associated with rainfall amounts. 
```

Making new figure

```{r}
ggplot(eastwood, aes(Rainfall.nestling, rTL))+
     geom_point()+
     geom_smooth(method = "lm")+
     labs(x = "Rainfall during nestling phase", y = "Telomere length")+
     theme_bw()
```

Results Rainfall amount during the nestling phase for superb fairy wrens does not impact telomere length (F~1,341~ = 0.1253, p = 0.7236).

#Question 2: Is there a relationship between minimum temperature and telomere length? I am looking for a relationship between temperature and telomere length, which are both ratio data. Since there is only one predictor variable, I will use a linear regression.

Predicting analysis

```{r}
#Making scatterplot as initial depiction of data. Making predictions off of this scatterplot can allow me to determine if there are any mistakes when I do my analysis. The summary statistics in the analysis should be similar to the ones estimated from the scatterplot.
ggplot(eastwood, aes(Tmin.nestling, rTL)) + geom_point()

#The telomere lengths range between 0.25-1.20, and the temperatures range between 4-18. The slope should be about (1.2-.25)/(18-4) = 0.07. The y-intercept appears to be around the origin. Because of this low slope, I do not expect there to be a significant relationship between telomere length and minimum temperature. Also, there does not appear to be a relationship based just upon a visual evaluation of the graph.
```

Making linear regression

```{r}
model_temperature <- lm(rTL ~ Tmin.nestling, data = eastwood)
```

Checking assumption

```{r}
#The residuals vs fitted plot should have no humps or valleys to assume that a linear model is appropriate for this data. In the Q-Q plot, the data should be linear to assume that the data is normally distributed. In the scale-location plot, there should be no pattern of points, so the line should be straight, to assume that there is constant variance. The residuals vs leverage plot detects points that have too much influence on the data set, such as outliers. 
autoplot(model_temperature, smooth.color = NA)

#The residuals vs fitted plot and Q-Q plot look good. The scale-location and residuals vs leverage plots both have a slight positive slope, but they appear mild enough for the data to still fit the assumptions of this model.
```

Running ANOVA

```{r}
#ANOVA test
anova(model_temperature)
#The F-statistic is 0.8925 and the p-value is 0.3455, which does not indicate a significant relationship between minimum temperature and telomere length. The degrees of freedom + residuals + 1 should add up to the number of rows, which is the case because 1 + 341 + 1 = 343. 

#Viewing results in a summary table
summary(model_temperature)
#The adjusted R-squared is -0.0003143. This value indicates that the variation in telomere length is not highly associated with temperature. 
```

Making new figure

```{r}
ggplot(eastwood, aes(Tmin.nestling, rTL))+
     geom_point()+
     geom_smooth(method = "lm")+
     labs(x = "Minimum temperature during nestling phase", y = "Telomere length")+
     theme_bw()
```

Results Minimum temperature during the nestling phase for superb fairy wrens does not impact telomere length (F~1,341~ = 0.8925, p = 0.3455).
